\documentclass[11pt]{article}

\usepackage[a4paper, margin=1in]{geometry}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{hyperref}
\usepackage{titlesec}
\usepackage{enumitem}

\titleformat{\section}{\large\bfseries}{\thesection}{1em}{}
\titleformat{\subsection}{\normalsize\bfseries}{\thesubsection}{1em}{}

\title{\textbf{CP260-2026 Final Project Report}\
\large Metric-Semantic Reconstruction}

\author{
Shivam Hake (27141) \
Aryan Dahiya (26579)
}

\date{}

\begin{document}

\maketitle

\section{Problem Description}

The objective of this project is to perform \textbf{metric-semantic reconstruction} of a scene using RGB images and corresponding camera poses. The system reconstructs a 3D scene while identifying specific objects and estimating their \textbf{oriented bounding boxes (OBB)} in metric scale.

The primary challenges include:
\begin{itemize}
\item Accurate 3D reconstruction from sparse multi-view images
\item Semantic identification of objects using natural language queries
\item Maintaining real-world scale consistency
\end{itemize}

The system is evaluated using \textbf{Intersection over Union (IoU)} on unseen test images.

\section{Literature Review}

Recent advancements combine geometric reconstruction with deep learning:

\subsection{Structure-from-Motion (SfM)}
Classical methods reconstruct 3D scenes but lack semantic understanding.

\subsection{Depth Estimation}
Models like \textbf{Depth Anything V2} estimate metric depth from single images.

\subsection{Open-Vocabulary Detection}
\textbf{Grounding DINO} enables text-conditioned object detection.

\subsection{Segmentation}
\textbf{Segment Anything Model (SAM)} provides high-quality object masks.

\subsection{3D Bounding Boxes}
PCA-based methods estimate orientation and extent from point clouds.

\section{System Architecture}

\subsection{Pipeline Overview}

\begin{center}
Input Images + Poses $\rightarrow$ Depth Estimation $\rightarrow$ 3D Reconstruction $\rightarrow$ Detection $\rightarrow$ Segmentation $\rightarrow$ 3D Point Extraction $\rightarrow$ OBB Fitting $\rightarrow$ Output
\end{center}

\subsection{Design Choices}

\begin{itemize}
\item \textbf{Depth Anything}: for metric depth estimation
\item \textbf{Grounding DINO}: for flexible text-based detection
\item \textbf{SAM}: for accurate segmentation
\item \textbf{Multi-frame fusion}: improves robustness
\item \textbf{Scale correction}: ensures real-world accuracy
\end{itemize}

\section{Implementation Details}

\subsection{Data Processing}
Images and poses are loaded and aligned using frame IDs.

\subsection{Depth Estimation}
Depth maps are generated per image and scaled using known object dimensions.

\subsection{3D Reconstruction}
Depth is projected into 3D using camera intrinsics and poses.

\subsection{Detection and Segmentation}
Objects are detected using text prompts and segmented using SAM.

\subsection{Point Aggregation}
Points belonging to each object are collected across multiple frames.

\subsection{OBB Fitting}
Principal Component Analysis (PCA) is used to compute oriented bounding boxes.

\subsection{Challenges Addressed}
\begin{itemize}
\item Frame–pose mismatch resolution
\item API compatibility issues with Grounding DINO
\item Depth scale correction
\item Handling sparse datasets
\end{itemize}

\section{Experimental Results}

\subsection{Qualitative Results}
The system successfully detects:
\begin{itemize}
\item Ethernet socket
\item Power socket
\item HDMI socket (left)
\item USB socket (top-right)
\end{itemize}

\subsection{Quantitative Results}

\begin{itemize}
\item IoU: approximately 0.6 -- 0.8
\item Improved accuracy after scale correction
\item Better detection with refined prompts
\end{itemize}

\subsection{Observations}
\begin{itemize}
\item Prompt design significantly impacts detection quality
\item Scale correction improves metric accuracy
\item Multi-view fusion reduces noise
\end{itemize}

\section{Conclusions}

This project demonstrates an effective pipeline for combining geometric reconstruction with semantic understanding.

Key contributions:
\begin{itemize}
\item Integration of depth estimation and semantic detection
\item Accurate metric-scale reconstruction
\item Robust multi-view fusion
\end{itemize}

\section{References}

\begin{enumerate}
\item Depth Anything V2
\item Grounding DINO
\item Segment Anything Model (SAM)
\item COLMAP (SfM framework)
\item PCA-based 3D bounding box estimation
\end{enumerate}

\section{Bonus: Additional Experiments}

\begin{itemize}
\item Prompt engineering for improved detection
\item Forced depth scaling for better IoU
\item Context-aware queries (e.g., relative to VGA port)
\item Optimization for sparse input frames
\end{itemize}

\end{document}
